Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-19 main.tex
minor fixes

For a formula $A$, we can define its
associated fograph $\graphof{A}$ inductively by:
\begin{equation*}
  \begin{array}{cc}
    \graphof{\alpha} = \single \alpha & 
    \graphof{\forall x . A} = \single x + \graphof{A} \\[.5ex]
    \graphof{A \cor B} = \graphof{A} + \graphof{B} &
    \graphof{\exists x . A} = \single x \times \graphof{A} \\[.5ex]
    \graphof{A \cand B} = \graphof{A} \times \graphof{B} &

  \end{array}
\end{equation*}
where $\single \alpha$ (resp. $\single x$) is a single-vertex graph whose vertex is 
    labelled by $\alpha$ (resp. $x$).
\todo{make look nice, use array} \juihsuan{done}\lutz{use the single macro (see
above), so that we can change the notation later if we need to (and make the
array with two columns)} \juihsuan{fixed}

@@

\todo{def $\fequ$}

We define a congruence relation $\fequ$ on formulas by the following equations:
\begin{equation}
  \label{eq:fequ}
  \begin{array}{cc}
    A \vlan B \fequ B \vlan A & A \vlor B \fequ B \vlor A \\
    (A \vlan B) \vlan C \fequ A \vlan (B \vlan C) &
    (A \vlor B) \vlor C \fequ A \vlor (B \vlor C) \\
    \forall x . \forall y . A \fequ \forall y . \forall x . A &
    \exists x . \exists y . A \fequ \exists y . \exists x . A \\
    \forall x . (A \vlor B) \fequ (\forall x . A) \vlor B &
    \exists x . (A \vlan B) \fequ (\exists x . A) \vlan B
  \end{array}
\end{equation}
where $x \notin fv(B)$ in the last two equations.
Two formulas $A$ and $B$ are \bfit{equivalent} if $A \fequ B$. The following
theorem shows that the set of fographs can been seen as the quotient
$\FORM / \fequ$.
\begin{thm}
  Let $A,B$ be rectified formulas. Then
  $$
  A\fequ B \iff \graphof A =\graphof B
  $$
\end{thm}

\begin{proof}
  \todo{}
  By a straightforward induction on $A$.
\end{proof}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

          \vlhy{\Scons{(R_1\cons{x}\vlan C)\vlor (R_2\cons{x}\vlor D)}}}}}

@@

              \vlhy{\Scons{((x\vlor E)\vlan C)\vlor ((x\vlor F)\vlor D)}}}}}

@@

              \vlhy{\Scons{((\forall x. E)\vlan C)\vlor ((\forall x. F)\vlor D)}}}}}

@@

is a propositional variable.}

@@

          \vlhy{\Scons{(x\vlan C)\vlor ((x\vlor F)\vlor D)}}}}}

@@

              \vlhy{\Scons{(x\vlan C)\vlor (x\vlor D)}}}}}

>>>>

For a formula $A$, we can define its
associated fograph $\graphof{A}$ inductively by:
\begin{equation*}
  \begin{array}{r@{\;}l@{\hskip4em}r@{\;}l}
    \graphof{a} &= \single a \rlap{\hskip1em(for any atom $a$)}  \\[.9ex]
    \graphof{A \cor B} &= \graphof{A} + \graphof{B} &
    \graphof{\exists x . A} &= \single x \times \graphof{A} \\[.9ex]
    \graphof{A \cand B} &= \graphof{A} \times \graphof{B} &
    \graphof{\forall x . A} &= \single x + \graphof{A}    
  \end{array}
\end{equation*}
where $\single a$ (resp. $\single x$) is a single-vertex graph whose vertex is 
    labelled by $a$ (resp. $x$).

@@

We define a congruence relation $\fequ$ on formulas by the following equations:
\begin{equation}
  \label{eq:fequ}
  \begin{array}{c@{\hskip1.5em}c}
    A \vlan B \fequ B \vlan A
    &
    (A \vlan B) \vlan C \fequ A \vlan (B \vlan C) 
    \\
    A \vlor B \fequ B \vlor A
    &
    (A \vlor B) \vlor C \fequ A \vlor (B \vlor C)
    \\
    \forall x . \forall y . A \fequ \forall y . \forall x . A
    &
    \forall x . (A \vlor B) \fequ (\forall x . A) \vlor B
    \\
    \exists x . \exists y . A \fequ \exists y . \exists x . A
    &
    \exists x . (A \vlan B) \fequ (\exists x . A) \vlan B
  \end{array}
\end{equation}
where $x \notin fv(B)$ in the last two equations.
Two formulas $A$ and $B$ are \bfit{equivalent} if $A \fequ B$. The following
theorem shows that the set of fographs can been seen as the quotient
$\FORM/\mathord\fequ$.
\begin{thm}
  Let $A,B$ be rectified formulas. Then
  $$
  A\fequ B \iff \graphof A =\graphof B
  $$
\end{thm}

\begin{proof}
  By a straightforward induction on $A$.
\end{proof}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

          \vlhy{\Scons{(R_1\cons{x}\vlan C)\vlor (R_2\cons{x}\vlan D)}}}}}

@@

              \vlhy{\Scons{((x\vlor E)\vlan C)\vlor ((x\vlor F)\vlan D)}}}}}

@@

              \vlhy{\Scons{((\forall x. E)\vlan C)\vlor ((\forall x. F)\vlan D)}}}}}

@@

is a propositional variable.}\lutz{shit. (you mean a ``first order variable'') this actually can happen. then we have another $\mexD$}

@@

          \vlhy{\Scons{(x\vlan C)\vlor ((x\vlor F)\vlan D)}}}}}

@@

              \vlhy{\Scons{(x\vlan C)\vlor (x\vlan D)}}}}}
